---
title: "Social Network Analysis - Exercise 2"
author: PairJS - Javier Granda - Youssef Sami
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

On this first exercise, we will learn the basic functions of the Igraph package, a great tool to use Social Network Analysis metrics in our projects. We will use in this first exercise a dataset, "mathoverflow-ints.txt", containing interactions between MathOverflow users with the next information:

- Answers to questions
- Comments to questions
- Comments to answers

You have to complete the code chunks in this document but also analyze the results, extract insights and describe them in an understandable way.

<span style="color:blue">Please, write all your conclusions in blue</span>

Both exercises may be done individually or in pairs; do not forget to write the name of the members of the team at the beginning of this document and also your team number in the name of this file (exercise1_PXX.Rmd)

```{r echo=T, message=FALSE, warning=FALSE}

library(knitr)
library (dplyr)
library(ggplot2)
library(igraph)
```


## Loading data

In this section, the goal is loading the dataset given for this exercise, building the graph object and analyzing basics metrics.

- Build the graph object with all the interactions. Include the timestamp and the type of the edge as edge attributes.
- Describe the values provided by summary function on the graph object.
- Validate the graph has been created correctly comparing with the information after reading the file.
- Simplify the graph in order to remove loops and multiple edges. After simplifying the graph, a weight attribute should have been created, counting the number of interactions between two users.

# Question 1

### Build the graph object with all the interactions. Include the timestamp and the type of the edge as edge attributes
```{r}
# CHUNK 1
# function that transform all numeric columns into character to transform the nodes number as character
to_char<-function(df){
  for(i in 1:ncol(df))
{
  if(is.numeric(df[[i]]))
  {
    df[[i]]<- as.character( df[[i]])
  }
  
  }
  return (df)
}
# reading the data file 
SN<-data.frame(read.table('mathoverflow-ints.txt'))
# remove the link type form the data frame to analyze all the type of links
SN<-SN[,1:3]
SN<-to_char(SN)

# create an igraph instance from the social network
igraph <- graph.data.frame(SN, directed=FALSE)
E(igraph)$W
```
### Describe the values provided by summary function on the graph object

The ighraph summary:
First Line:
first is the type of the object IGRAPH
UNW Undirected, Named graph and the 24818 as the number of Edges and 506550 number of nodes
Second Line:
Igraph attributes: Vertex, and edge attribute V3

```{r}
summary(igraph)
```

### Validate the graph has been created correctly comparing with the information after reading the file
```{r}
# get the igraph nodes in each edge in a datafarame 
edgeList<-to_char(data.frame(get.edgelist(igraph)))
# create a vector with all the unique nodes from the read file
nodes<-c(unique(SN$V1),unique(SN$V2))
# compare the number of nodes from the read file and the nodes from the Igraph. if equals return true
vcount(igraph)==length(unique(nodes))
# compare the number of edges
nrow(edgeList)==nrow(SN)
# check if the timestamps in igraph equals the timestamps in the read file
sum(E(igraph)$V3!=SN$V3)
# check if each edge has the same nodes between the created igraph and the read file
sum(edgeList[i,1:2] != SN[i, 1:2] & edgeList[i,1:2] != SN[i, 2:1])
```



### Simplify the graph in order to remove loops and multiple edges. After simplifying the graph, a weight attribute should have been created, counting the number of interactions between two users.
```{r}
# Create and edge attribute of weight equals to 1
E(igraph)$weight=1
# simplify the igraph to remove loops and duplicates edges
SimpleG<-simplify(igraph)
```



# Question 2:

## Analyzing the graph


### - Is the graph connected? If it is not, compute the connected components and analyze the statistical distribution of the size of the resulting components. Do this considering all the links but also using only the different types of edges.



```{r}
is.connected(SimpleG)
#Get number of cluster
no.clusters(SimpleG)
clu <- clusters(SimpleG)
# get different groups
groups<-groups(clu)

graphs <- decompose.graph(SimpleG,min.vertices = 1)
#Clusters sizes distributions 
sizeDist<-as.data.frame(table(clu$csize))


ggplot(sizeDist, aes(x=Var1, y=Freq)) + geom_bar(stat="identity", fill="steelblue")


```


### Analyze the statistical distribution of the "weight" attribute created in the previous section and fit it to a power law distribution. Does it fit well enough? 
```{r}

G.weights <- E(SimpleG)$weight
weightsDist<-as.data.frame(table(G.weights))
weightsDist$G.weights<-as.numeric(as.character(weightsDist$G.weights))

ggplot(weightsDist, aes(x=G.weights, y=Freq)) + geom_bar(stat="identity", fill="steelblue") 
ggplot(weightsDist, aes(x=G.weights, y=Freq)) +
  geom_point() +
  scale_x_continuous("weight\n(edge with this amount of connections)",
                     breaks = c(1, 3, 10, 30, 100, 300),
                     trans = "log10") +
  scale_y_continuous("Frequency\n(how many of them)",
                     breaks = c(1, 3, 10, 30, 100, 300, 1000),
                     trans = "log10") +
  ggtitle("Weights Distribution (log-log)") +
theme_bw()
```



## Node degree

A typical analysis when dealing with social networks is studying the degree distribution of the nodes.
- Visualize the statisical distribution of the node degree.
- Again, is it a power-law distribution? Explain why this is the case.

```{r}
# CHUNK 3
G.degrees <- degree(SimpleG)

degreeDist <- as.data.frame(table(G.degrees))
degreeDist$G.degrees<-as.numeric(as.character(degreeDist$G.degrees))

ggplot(degreeDist, aes(x=G.degrees, y=Freq)) + geom_bar(stat="identity", fill="steelblue") 

ggplot(degreeDist, aes(x = G.degrees, y = Freq)) +
  geom_point() +
  scale_x_continuous("Degree\n(nodes with this amount of connections)",
                     trans = "log10") +
  scale_y_continuous("Frequency\n(how many of them)",
                     trans = "log10") +
  ggtitle("Degree Distribution (log-log)") +
theme_bw()
```

## Building connected graphs

In general, in interaction networks, many connected components turn out after creating the graph. However, most of the nodes typically belong to the same giant component. In this section, the goal is:
- Build a subgraph with all the nodes belonging to the largest connected component.
- Validate the graph has been correctly created.
- Analyze how node degree has changed from the original graph to this new subgraph. Find out which nodes have now a smaller number of connections and explain why.


```{r}
# CHUNK 4
largest <- which.max(sapply(graphs, vcount))
plot(graphs[[largest]], layout=layout.fruchterman.reingold)


plot(g, layout=layout.fruchterman.reingold)
```

# Visualizing shortest paths

On this previous subgraph, you have to compute:

- The shortest path from the node with the highest degree to the rest of the graph.
- From all these paths, select one path with the maximum length.
- Visualize only this path using the plot function.
- Visualize this path within the giant component subgraph.

To this end, use the layout.fruchterman.reingold function to place the nodes in the visualization.

```{r}
# CHUNK 5
```